---
title: "Lamb_pain_Cortisol_all"
author: "Shuaimin Kang"
date: "8/8/2017"
output: html_document
---

```{r echo=FALSE}
setwd("/Users/skang/Downloads/2017Consulting/lamb MEL")
load("KBeltaireStatsNumeric.RData")
library("ggplot2")
library("tidyr")
Cort <- gather(C.all1, value = "value",
          key = "type", `-0.5`, `0.25`, `0.75`, `3`, `12`, `24`)
Cort$type <- as.numeric(Cort$type)
```

#Plot time vs cortisol level by person by group
```{r echo=FALSE}
ggplot(Cort, aes(x=type, y=value,color=Lamb)) + geom_point()+ geom_line() + facet_wrap(~Tx, ncol = 10) + labs(x="time",y="cortisol")

```

#Plot time vs increase of cortisol level since initial time -0.5 by person by group
```{r echo=FALSE}
Cort1 <- Cort
for (i in 1:length(Cort1$Lamb)){
  Cort1$value[i] <- Cort1$value[i] - Cort$value[Cort1$type==-0.5 & Cort1$Lamb==Cort1$Lamb[i]]
}

ggplot(Cort1, aes(x=type, y=value,color=Lamb)) + geom_point()+ geom_line() + facet_wrap(~Tx, ncol = 10) + labs(x="time",y="cortisol")
```

```{r echo=FALSE}

gm <- aggregate(Cort["value"],list(Cort$Tx,Cort$type),FUN=mean)
gm1 <- gm
gm1$value[gm1$Group.1=="C"] = gm1$value[gm1$Group.1=="C"] - gm1$value[gm1$Group.1=="C"][1]

gm1$value[gm1$Group.1=="M"] = gm1$value[gm1$Group.1=="M"] - gm1$value[gm1$Group.1=="M"][1]

names(gm1) <- c("Group","time","cortisol")
```

#Plot of difference of mean cortisol level at time t and at initial time t=-0.5. 

```{r echo=FALSE}
ggplot(gm1, aes(x=time, y=cortisol,color=Group)) + geom_point() + geom_line() + ggtitle("mean cortisol level at t - mean cortisol level at t=-0.5")
```

```{r echo=FALSE}
ttestf <- function(dat, time, alternative){
  dat.Control <- dat$value[dat$Tx=="C" & dat$type==time] - dat$value[dat$Tx=="C" & dat$type==-0.5]
  dat.Treat <- dat$value[dat$Tx=="M" & dat$type==time]- dat$value[dat$Tx=="M" & dat$type==-0.5]
  return(t.test(dat.Control, dat.Treat, alternative=alternative))
}


wilcoxtestf <- function(dat, time, alternative){
  dat.Control <- dat$value[dat$Tx=="C" & dat$type==time] - dat$value[dat$Tx=="C" & dat$type==-0.5]
  dat.Treat <- dat$value[dat$Tx=="M" & dat$type==time]- dat$value[dat$Tx=="M" & dat$type==-0.5]
  return(wilcox.test(dat.Control-dat.Treat, alternative=alternative))
}
```


t-test for mean difference: cortisol level at time t=0.25 - at initial time t=-0.5
```{r echo=FALSE}
print(paste0("t-test for Cortosis Level at time ", 0.25))
print(ttestf(dat=Cort, time= 0.25, alternative = "greater"))
```


t-test for mean difference: cortisol level at time t=0.75 - at initial time t=-0.5

```{r echo=FALSE}
print(paste0("t-test for Cortosis Level at time ", 0.75))
print(ttestf(dat=Cort, time= 0.75, alternative = "greater"))
```


t-test for mean difference: cortisol level at time t=3 - at initial time t=-0.5
```{r echo=FALSE}
print(paste0("t-test for Cortosis Level at time ", 3))
print(ttestf(dat=Cort, time= 3, alternative = "greater"))
```

t-test for mean difference: cortisol level at time t=12 - at initial time t=-0.5

```{r echo=FALSE}
print(paste0("t-test for Cortosis Level at time ", 12))
print(ttestf(dat=Cort, time= 12, alternative = "less"))
```

t-test for mean difference: cortisol level at time t=24 - at initial time t=-0.5

```{r echo=FALSE}
print(paste0("t-test for Cortosis Level at time ", 24))
print(ttestf(dat=Cort, time= 24, alternative = "less"))
```
# Mixed effect and fixed effect model on increase of cortisol level, order = 2


```{r echo=FALSE}
library(lme4)
#print(Cort[Cort$Lamb==514,])
#print(Cort1[Cort1$Lamb==514,])
Cort <- Cort1
#print(Cort1)

Cort$t <- Cort$type
Cort$t2 <- (Cort$type)^2

lmm1 <- lmer(value ~ (t+t2)*Tx + (1 | Lamb), data = Cort, REML=FALSE)
lmm2 <- lmer(value ~ (t+t2)*Tx + (1 + t| Lamb), data = Cort, REML=FALSE)
lmm3 <- lmer(value ~ (t+t2)*Tx + (1 + t + t2| Lamb), data = Cort, REML=FALSE)

lm <- lm(value ~ (t+t2)*Tx, data=Cort)

anova(lmm1, lmm2, lmm3)
summary(lmm1)
lmm11 <- lmer(value ~ t + t2 + Tx + t*Tx + (1 | Lamb), data = Cort, REML=FALSE)
lmm12 <- lmer(value ~ t + t2 + Tx + (1 | Lamb), data = Cort, REML=FALSE)

anova(lmm12, lmm11, lmm1)
summary(lmm11)
#summary(lmm3)
#summary(lm)

```


# Mixed effect and fixed effect model on increase of cortisol level, order = 1 Linear
```{r echo=FALSE}
library(lme4)
Cort$t <- Cort$type
Cort$t2 <- (Cort$type)^2

lmm1 <- lmer(value ~ (t)*Tx + (1 | Lamb), data = Cort, REML=FALSE)
lmm2 <- lmer(value ~ (t)*Tx + (1 + t| Lamb), data = Cort, REML=FALSE)

lm <- lm(value ~ (t)*Tx, data=Cort)

anova(lmm2, lmm1, lm)
summary(lmm1)
lmm11 <- lmer(value ~ t + Tx + (1 | Lamb), data = Cort, REML=FALSE)

anova(lmm1, lmm11)
```

### linear slope for C: -0.097, linear slope for M: -0.1+0.0768=-0.063;
Cortisol level is more stable for lambs taking MEL.   